---
title: "crab_compare"
output: html_document
date: "2026-05-17"
---

# Pull tables

```{r data_dl}

files <-  c("crab_sizecomp", "crab_biomass", "crab_cpue", "crab_specimen_preinclude", 
            "crab_specimen_pack", "crab_specimen", "hauls_retow",
            "crab_sizecomp_pack", "crab_biomass_pack", "crab_cpue_pack")
for (filesi in files) {
  temp <- read.csv(file = here::here(paste0("data/", filesi, ".csv"))) |> 
    dplyr::rename_all(tolower)
  assign(x = filesi, temp)
}

gapindex_sizecomp <- crab_sizecomp |> 
  dplyr::filter(area_id %in% c(99900, 99902)) |>
  dplyr::select(-area_id) |>
  dplyr::filter(!is.na(population_count)) |>
  dplyr::filter(population_count != 0)
gapindex_biomass <- crab_biomass
gapindex_cpue <- crab_cpue  |> 
  dplyr::filter(weight_kg > 0) |> 
  dplyr::mutate(
    area_swept_km2 = round(x = area_swept_km2, digits = 4))
crabpack_sizecomp <- crab_sizecomp_pack |> 
  dplyr::mutate(population_count = round(x = population_count, digits = 0)) |> # can't have fractions of organisms
  dplyr::filter(population_count != 0) |> 
  dplyr::filter(!is.na(population_count))
crabpack_biomass <- crab_biomass_pack
crabpack_cpue <- crab_cpue_pack |> 
  dplyr::filter(weight_kg > 0) |> 
  dplyr::mutate(
    area_swept_km2 = round(x = area_swept_km2, digits = 4)) 
```

## Catch per unit effort

```{r compare-cpue}
cpue_compare <- dplyr::full_join(
  crabpack_cpue, # |> dplyr::mutate(crabpack = 1), 
  gapindex_cpue, # |> dplyr::mutate(gapindex = 1), 
  by = c("hauljoin", "species_code")) # |> 
  # dplyr::mutate(source = gapindex+crabpack)
  setdiff(crabpack_cpue$hauljoin, gapindex_cpue$hauljoin)
  setdiff(gapindex_cpue$hauljoin, crabpack_cpue$hauljoin)
```
It is my understanding that none of the hauljoins in the crabpack data should be there. 

As I understand the retow data rule: haul_type 3 should be used except in retow years. In retow years, haul_type 17 data for female (2) RKC (69322) replaces any data for female RKC at those stations where haul_type 3. 

Firstly, there are haul_types that are not 3 and 17 in the crabpack data pull. These should not be used:
example code: `a <- crabpack_cpue |> dplyr::right_join(hauls_missing) |> dplyr::filter(haul_type == 13) ; unique(a$hauljoin)`
- haul_type = 20: Catch Selectively Sampled/Processed; 1 instance in 2006 (`hauljoin -11364`)
- haul_type = 13:	Index Sample Tow; several instances all form 2018 (`hauljoin -18178 -18188 -18189 -18185 -18184 -18145 -18191 -18210 -18197 -18196 -18180 -18195 -18176 -18177 -18192 -18208 -18211 -18173 -18174 -18179 -18193 -18206 -18207 -18212 -18219 -18216 -18220 -18221 -18222 -18190 -18194 -18205 -18204 -18213 -18214 -18217 -18218 -18203 -18202 -18181 -18183 -18144 -18209 -18187 -18215 -18175 -18223`)
- haul_type = 4: Fishing Power Comparative Sample; several instances all from 1982 (`hauljoin -4376 4386 4377 4385 4378 4384 4379 4383`)

Then, the haul_type 3 observation included in the crabpack pull but not in the gapindex data is for 3 female RKC (69322) (hauljoin -4811) that should be replaced with retow station data (-5757). Right now these are both being used in the estimates calculation. 

Then, the haul_type 17 included in the crabpack pull but not in the gapindex data
- Some observations are for male (1) RKC (69322) that should not be included, and should instead be represented by haul_type 3 observations (hauljoins: 965753  965736  965735  965733 1023485 1023470  -11704   -3821   -4443   -5703   -5690   -5691   -5701   -9379   -9378   -9374  -20687; `a <- hauls_missing_specimen |> dplyr::filter(dup == 1, include == 0, haul_type == 17) ; unique(a$hauljoin)`)
- Then there is one hauljoin (1023479) (and this station may be up for discussion) where there was no haul_type 3 trawl done at this station (E-11), only a haul_type 17 in 2000 EBS. Should this station/hauljoin be included, or excluded?


```{r}
  source("Z:/Projects/ConnectToOracle.R")
  hauls_missing <- RODBC::sqlQuery(channel = channel,
                         query = paste0("SELECT *FROM RACEBASE.HAUL WHERE HAULJOIN IN (",
                                        paste0(setdiff(crabpack_cpue$hauljoin, gapindex_cpue$hauljoin), collapse = ", "),
                                        ") ; ")) |> 
    dplyr::arrange(desc(HAUL_TYPE)) |> 
    dplyr::select(CRUISEJOIN, CRUISE, HAULJOIN, VESSEL, HAUL_TYPE, STATION = STATIONID, STRATUM, PERFORMANCE, ABUNDANCE_HAUL) |> 
                       dplyr::rename_all(tolower)
  hauls_missing
  
  hauls_missing_cpue <- crabpack_cpue |> 
    dplyr::filter(hauljoin %in% setdiff(crabpack_cpue$hauljoin, gapindex_cpue$hauljoin)) |> 
    dplyr::full_join(hauls_missing) |> 
    dplyr::arrange(desc(haul_type)) |> 
    dplyr::select(-area_swept, -vessel)
  
  hauls_missing_specimen <- crab_specimen_preinclude |> 
    dplyr::select(hauljoin, haul_type, species_code, sex, sex0, length_mm, count, weight_g, year, key, dup, include) |> 
    dplyr::filter(hauljoin %in% setdiff(crabpack_cpue$hauljoin, gapindex_cpue$hauljoin))
```


```{r}
cpue_compare_diff <- cpue_compare |> 
  dplyr::mutate(
    weight_kg_diff = weight_kg.x - weight_kg.y,
    count_diff = count.x - count.y, 
    area_swept_km2_diff = area_swept_km2.x - area_swept_km2.y,
    count_diff = count.x - count.y, 
    cpue_kgkm2_diff = cpue_kgkm2.x - cpue_kgkm2.y, 
    cpue_nokm2_diff = cpue_nokm2.x - cpue_nokm2.y) |> 
  dplyr::select(hauljoin, species_code, dplyr::ends_with("_diff")) |> 
  dplyr::mutate(dplyr::across(dplyr::ends_with("_diff"), ~ ifelse(.x > 0.001, 1, 0)), 
                noteable_diff = (#weight_kg_diff  + area_swept_km2_diff + 
                  count_diff + cpue_kgkm2_diff + cpue_nokm2_diff)) |> 
  dplyr::filter(noteable_diff > 0)

cpue_compare_diff_data <- cpue_compare |> 
  dplyr::right_join(cpue_compare_diff |> 
                      dplyr::select(hauljoin, species_code, noteable_diff))

a <- cpue_compare_diff |>
  dplyr::summarize(
    weight_kg_diff = sum(weight_kg_diff, na.rm = TRUE),
    count_diff = sum(count_diff, na.rm = TRUE),
    area_swept_km2_diff = sum(area_swept_km2_diff, na.rm = TRUE),
    cpue_kgkm2_diff = sum(cpue_kgkm2_diff, na.rm = TRUE),
    cpue_nokm2_diff = sum(cpue_nokm2_diff, na.rm = TRUE)
  )

print(paste0("Of ", nrow(cpue_compare), " records ( gapindex: ",nrow(gapindex_cpue), ", crabpack ", nrow(crabpack_cpue), "), ", nrow(cpue_compare_diff_data), " did not match. These issues occured in: ", paste0(names(a), ": ", a, collapse = ", ")))

summary(cpue_compare_diff_data)

table(cpue_compare_diff_data$species_code, cpue_compare_diff_data$hauljoin)
write.csv(x = cpue_compare_diff_data, file = here::here("data/cpue_compare_diff_data.csv"))

```
Let's look at the data broken out:

It looks like Male RKC seem to be removed in the crabpack data CPUE calculations in retow years. I thought just Female RKC should be replaced during retows. Further, none of these hauls appear to have been retowed (note that all hauls listed here are haul_type = 3 and `dup` = 0 where if dup = 1 would indicate a retow of that station occurred). 

# Specify retow stations and years for BBRKC
## Retow years: 1999 2000 2006 2007 2008 2009 2010 2011 2012 2017 2021

```{r breakitout-cpue}
print("here are the differences between the resultant datasets")
cpue_compare_diff_data

print("for the hauljoins that had differences, here are the data broken out by sex. ")
crab_specimen_preinclude |> 
  dplyr::filter(hauljoin %in% cpue_compare_diff_data$hauljoin) 


print("All hauls with differences to see if these are retow issues. ")
b <- hauls_retow |>
  dplyr::rename_all(tolower) |> 
  dplyr::left_join(crab_specimen_preinclude  |>
                     dplyr::rename_all(tolower) |> 
                     dplyr::group_by(hauljoin, species_code, haul_type, sex, key, station_id, year, include, dup) |>
                     dplyr::summarise(weight_kg = sum(weight_g, na.rm = TRUE)/1000,
                                      count = sum(frequency, na.rm = TRUE)) |>
                     dplyr::ungroup()) |> 
  dplyr::select(haul_type, key, hauljoin, dup, station_id, year) |> 
  dplyr::filter(hauljoin %in% unique(cpue_compare_diff_data$hauljoin)) # |>
# summary()
b
# all of these years are retow years
sort(unique(b$year))
```

One example of the above in action: 


```{r cpue-deep-dive}
crab_specimen_preinclude  |>
  dplyr::rename_all(tolower) |> 
  dplyr::group_by(hauljoin, species_code, haul_type, sex, key, station_id, year, include, dup) |>
  dplyr::summarise(weight_kg = sum(weight_g, na.rm = TRUE)/1000,
                   count = sum(frequency, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::filter(hauljoin == -19922) #key == "I-092021EBS") 

cpue_compare_diff_data |>
  dplyr::filter(hauljoin == -19922)
```

## Biomass

```{r compare-biomass-abundance}
gapindex_biomass <- gapindex_biomass |> 
  dplyr::filter(area_id %in% c(99900, #99901, 
                               99902))
biomass_compare <- dplyr::full_join(
  gapindex_biomass |> 
    dplyr::select(species_code, year, survey_definition_id,
                  biomass_mt, population_count),
  crabpack_biomass  |> 
    dplyr::select(species_code, year, survey_definition_id, 
                  biomass_mt, population_count), 
  by = c("survey_definition_id", "year", "species_code"))

biomass_compare_diff <- biomass_compare |> 
  dplyr::mutate(biomass_mt_diff = biomass_mt.x - biomass_mt.y, 
                population_count_diff = population_count.x - population_count.y) |> 
  dplyr::select(survey_definition_id, year, species_code, 
                dplyr::ends_with("_diff")) |> 
  dplyr::mutate(dplyr::across(dplyr::ends_with("_diff"), ~ ifelse(.x > 0.01, 1, 0)), 
                noteable_diff = (biomass_mt_diff + population_count_diff)) |> 
  dplyr::filter(noteable_diff > 0)

# biomass_compare_diff

biomass_compare_diff_data <- biomass_compare |> 
  dplyr::right_join(biomass_compare_diff |> 
                      dplyr::select(survey_definition_id, year, species_code, noteable_diff))

a <- biomass_compare_diff |>
  dplyr::summarize(
    biomass_mt_diff = sum(biomass_mt_diff, na.rm = TRUE),
    population_count_diff = sum(population_count_diff, na.rm = TRUE)
  )

print(paste0("Of ", nrow(biomass_compare), 
             " records ( gapindex: ",nrow(gapindex_biomass), 
             ", crabpack ", nrow(crabpack_biomass), "), ", nrow(biomass_compare_diff_data), 
             " did not match. These issues occured in: ", 
             paste0(names(a), ": ", a, collapse = ", ")))
biomass_compare_diff_data
summary(biomass_compare_diff_data)

table(biomass_compare_diff_data$species_code, biomass_compare_diff_data$year, biomass_compare_diff_data$survey_definition_id)
write.csv(x = biomass_compare_diff_data, file = here::here("data/biomass_compare_diff_data.csv"))

```

## Sizecomp

```{r compare-sizecomp}
sizecomp_compare <- dplyr::full_join(
  gapindex_sizecomp |> dplyr::mutate(gapindex = 1), 
  crabpack_sizecomp |> dplyr::mutate(crabpack = 1), 
  by = c("survey_definition_id", "year", "species_code", "length_mm", "sex")) |> 
  dplyr::mutate(source = (crabpack+gapindex))


sizecomp_compare_diff <- sizecomp_compare |> 
  dplyr::mutate(population_count_diff = (population_count.x - population_count.y), 
                percent_diff = (population_count.y - population_count.x)/population_count.x,
                noteable_diff = ifelse(population_count_diff > 0.001, 1, 0)) |> 
  dplyr::filter(noteable_diff > 0)

a <- sizecomp_compare_diff |>
  dplyr::summarize(
    population_count_diff = sum(population_count_diff, na.rm = TRUE)
  )

print(paste0("Of ", nrow(sizecomp_compare), 
             " records ( gapindex: ",nrow(gapindex_sizecomp), 
             ", crabpack ", nrow(crabpack_sizecomp), "), ", nrow(sizecomp_compare_diff), 
             " did not match. "))
# sizecomp_compare_diff
summary(sizecomp_compare_diff)
sizecomp_compare_diff_data <- sizecomp_compare |>
  dplyr::right_join(sizecomp_compare_diff |>
                      dplyr::select(survey_definition_id, year, species_code, length_mm, sex, noteable_diff))
# 
# sizecomp_compare_diff_data
# 
table(sizecomp_compare_diff_data$species_code, sizecomp_compare_diff_data$year, sizecomp_compare_diff_data$survey_definition_id)

write.csv(x = sizecomp_compare_diff_data, file = here::here("data/sizecomp_compare_diff_data.csv"))

```

Let's dive in to an example

Is the difference here just rounding error with the total area unit conversion?

```{r}
sizecomp_compare  |> 
  dplyr::rename_all(tolower) |>
  dplyr::filter(year == 2017 & species_code == 68560 & survey_definition_id  == 143) |> 
  dplyr::full_join(
    crab_specimen_preinclude |> 
      dplyr::rename_all(tolower) |> 
      dplyr::filter(year == 2017 & species_code == 68560 & region == 'NBS') |>
      dplyr::select(hauljoin, species_code, sex, length_mm, weight_g, haul_type, station_id) |> 
      dplyr::left_join(crabpack_cpue |> dplyr::select(-weight_kg, -count))
  )
#  |> 
# dplyr::mutate(
#   AREA_ID = ifelse(station_id == "AA-19", 71, 81), 
#   AREA_NM2 = ifelse(AREA_ID == 71, 57744 * x * 3.4299, # nmi2 to km2
#                     57744 * y * 3.4299, # nmi2 to km2
#                     ), # being used by crabpack
#   AREA_KM2 = ifelse(AREA_ID == 71, 81255.025, 38351.89), # from GAP_PRODUCTS
#   population_count.ytest = population_count.y/AREA_KM2
# )


```


