---
title: "crab_compare"
output: html_document
date: "2026-05-17"
---

# Pull tables

```{r data_dl}

files <-  c("crab_sizecomp", "crab_biomass", "crab_cpue", "crab_specimen_preinclude", 
            "crab_specimen_pack", "crab_specimen",
            "crab_sizecomp_pack", "crab_biomass_pack", "crab_cpue_pack")
for (filesi in files) {
  temp <- read.csv(file = here::here(paste0("data/", filesi, ".csv"))) |> 
    dplyr::rename_all(tolower)
  assign(x = filesi, temp)
}

gapindex_sizecomp <- crab_sizecomp |> 
  dplyr::filter(area_id %in% c(99900, 99902)) |>
  dplyr::select(-area_id) |>
  dplyr::filter(!is.na(population_count)) |>
  dplyr::filter(population_count != 0)
gapindex_biomass <- crab_biomass
gapindex_cpue <- crab_cpue  |> 
  dplyr::filter(weight_kg > 0) |> 
  dplyr::mutate(
    area_swept_km2 = round(x = area_swept_km2, digits = 4))
crabpack_sizecomp <- crab_sizecomp_pack |> 
  dplyr::mutate(population_count = round(x = population_count, digits = 0)) |> # can't have fractions of organisms
  dplyr::filter(population_count != 0) |> 
  dplyr::filter(!is.na(population_count))
crabpack_biomass <- crab_biomass_pack
crabpack_cpue <- crab_cpue_pack |> 
  dplyr::filter(weight_kg > 0) |> 
  dplyr::mutate(
    area_swept_km2 = round(x = area_swept_km2, digits = 4)) 
```

## Catch per unit effort

```{r compare-cpue}
cpue_compare <- dplyr::full_join(
  crabpack_cpue, # |> dplyr::mutate(crabpack = 1), 
  gapindex_cpue, # |> dplyr::mutate(gapindex = 1), 
  by = c("hauljoin", "species_code")) # |> 
  # dplyr::mutate(source = gapindex+crabpack)
  setdiff(crabpack_cpue$hauljoin, gapindex_cpue$hauljoin)
  setdiff(gapindex_cpue$hauljoin, crabpack_cpue$hauljoin)

cpue_compare_diff <- cpue_compare |> 
  dplyr::mutate(
    weight_kg_diff = weight_kg.x - weight_kg.y,
    count_diff = count.x - count.y, 
    area_swept_km2_diff = area_swept_km2.x - area_swept_km2.y,
    count_diff = count.x - count.y, 
    cpue_kgkm2_diff = cpue_kgkm2.x - cpue_kgkm2.y, 
    cpue_nokm2_diff = cpue_nokm2.x - cpue_nokm2.y) |> 
  dplyr::select(hauljoin, species_code, dplyr::ends_with("_diff")) |> 
  dplyr::mutate(dplyr::across(dplyr::ends_with("_diff"), ~ ifelse(.x > 0.001, 1, 0)), 
                noteable_diff = (#weight_kg_diff  + area_swept_km2_diff + 
                  count_diff + cpue_kgkm2_diff + cpue_nokm2_diff)) |> 
  dplyr::filter(noteable_diff > 0)

# cpue_compare_diff

a <- cpue_compare_diff |>
  dplyr::summarize(
    weight_kg_diff = sum(weight_kg_diff, na.rm = TRUE),
    count_diff = sum(count_diff, na.rm = TRUE),
    area_swept_km2_diff = sum(area_swept_km2_diff, na.rm = TRUE),
    cpue_kgkm2_diff = sum(cpue_kgkm2_diff, na.rm = TRUE),
    cpue_nokm2_diff = sum(cpue_nokm2_diff, na.rm = TRUE)
  )

cpue_compare_diff_data <- cpue_compare |> 
  dplyr::right_join(cpue_compare_diff |> 
                      dplyr::select(hauljoin, species_code, noteable_diff))

print(paste0("Of ", nrow(cpue_compare), " records ( gapindex: ",nrow(gapindex_cpue), ", crabpack ", nrow(crabpack_cpue), "), ", nrow(cpue_compare_diff_data), " did not match. These issues occured in: ", paste0(names(a), ": ", a, collapse = ", ")))

summary(cpue_compare_diff_data)

table(cpue_compare_diff_data$species_code, cpue_compare_diff_data$hauljoin)
write.csv(x = cpue_compare_diff_data, file = here::here("data/cpue_compare_diff_data.csv"))

```
Let's look at the data broken out:

It looks like Male RKC seem to be removed in the crabpack data CPUE calculations in retow years. I thought just Female RKC should be replaced during retows. Further, none of these hauls appear to have been retowed (note that all hauls listed here are haul_type = 3 and `dup` = 0 where if dup = 1 would indicate a retow of that station occurred). 

# Specify retow stations and years for BBRKC
## Retow years: 1999 2000 2006 2007 2008 2009 2010 2011 2012 2017 2021

```{r breakitout-cpue}
print("here are the differences between the resultant datasets")
cpue_compare_diff_data

print("for the hauljoins that had differences, here are the data broken out by sex. ")
crab_specimen_preinclude |> 
  dplyr::filter(hauljoin %in% cpue_compare_diff_data$hauljoin) 


print("All hauls with differences to see if these are retow issues. ")
hauls_retow <- read.csv(here::here("data/hauls_retow.csv"))
b <- hauls_retow |>
  dplyr::rename_all(tolower) |> 
  dplyr::left_join(crab_specimen_preinclude  |>
                     dplyr::rename_all(tolower) |> 
                     dplyr::group_by(hauljoin, species_code, haul_type, sex, key, station_id, year, include, dup) |>
                     dplyr::summarise(weight_kg = sum(weight_g, na.rm = TRUE)/1000,
                                      count = sum(frequency, na.rm = TRUE)) |>
                     dplyr::ungroup()) |> 
  dplyr::select(haul_type, key, hauljoin, dup, station_id, year) |> 
  dplyr::filter(hauljoin %in% unique(cpue_compare_diff_data$hauljoin)) # |>
# summary()
b
# all of these years are retow years
sort(unique(b$year))
```

One example of the above in action: 


```{r cpue-deep-dive}
crab_specimen_preinclude  |>
  dplyr::rename_all(tolower) |> 
  dplyr::group_by(hauljoin, species_code, haul_type, sex, key, station_id, year, include, dup) |>
  dplyr::summarise(weight_kg = sum(weight_g, na.rm = TRUE)/1000,
                   count = sum(frequency, na.rm = TRUE)) |>
  dplyr::ungroup() |>
  dplyr::filter(hauljoin == -19922) #key == "I-092021EBS") 

cpue_compare_diff_data |>
  dplyr::filter(hauljoin == -19922)
```

## Biomass

```{r compare-biomass-abundance}
gapindex_biomass <- gapindex_biomass |> 
  dplyr::filter(area_id %in% c(99900, #99901, 
                               99902))
biomass_compare <- dplyr::full_join(
  gapindex_biomass |> 
    dplyr::select(species_code, year, survey_definition_id,
                  biomass_mt, population_count),
  crabpack_biomass  |> 
    dplyr::select(species_code, year, survey_definition_id, 
                  biomass_mt, population_count), 
  by = c("survey_definition_id", "year", "species_code"))

biomass_compare_diff <- biomass_compare |> 
  dplyr::mutate(biomass_mt_diff = biomass_mt.x - biomass_mt.y, 
                population_count_diff = population_count.x - population_count.y) |> 
  dplyr::select(survey_definition_id, year, species_code, 
                dplyr::ends_with("_diff")) |> 
  dplyr::mutate(dplyr::across(dplyr::ends_with("_diff"), ~ ifelse(.x > 0.01, 1, 0)), 
                noteable_diff = (biomass_mt_diff + population_count_diff)) |> 
  dplyr::filter(noteable_diff > 0)

# biomass_compare_diff

biomass_compare_diff_data <- biomass_compare |> 
  dplyr::right_join(biomass_compare_diff |> 
                      dplyr::select(survey_definition_id, year, species_code, noteable_diff))

a <- biomass_compare_diff |>
  dplyr::summarize(
    biomass_mt_diff = sum(biomass_mt_diff, na.rm = TRUE),
    population_count_diff = sum(population_count_diff, na.rm = TRUE)
  )

print(paste0("Of ", nrow(biomass_compare), 
             " records ( gapindex: ",nrow(gapindex_biomass), 
             ", crabpack ", nrow(crabpack_biomass), "), ", nrow(biomass_compare_diff_data), 
             " did not match. These issues occured in: ", 
             paste0(names(a), ": ", a, collapse = ", ")))
biomass_compare_diff_data
summary(biomass_compare_diff_data)

table(biomass_compare_diff_data$species_code, biomass_compare_diff_data$year, biomass_compare_diff_data$survey_definition_id)
write.csv(x = biomass_compare_diff_data, file = here::here("data/biomass_compare_diff_data.csv"))

```

## Sizecomp

```{r compare-sizecomp}
sizecomp_compare <- dplyr::full_join(
  gapindex_sizecomp |> dplyr::mutate(gapindex = 1), 
  crabpack_sizecomp |> dplyr::mutate(crabpack = 1), 
  by = c("survey_definition_id", "year", "species_code", "length_mm", "sex")) |> 
  dplyr::mutate(source = (crabpack+gapindex))


sizecomp_compare_diff <- sizecomp_compare |> 
  dplyr::mutate(population_count_diff = (population_count.x - population_count.y), 
                percent_diff = (population_count.y - population_count.x)/population_count.x,
                noteable_diff = ifelse(population_count_diff > 0.001, 1, 0)) |> 
  dplyr::filter(noteable_diff > 0)

a <- sizecomp_compare_diff |>
  dplyr::summarize(
    population_count_diff = sum(population_count_diff, na.rm = TRUE)
  )

print(paste0("Of ", nrow(sizecomp_compare), 
             " records ( gapindex: ",nrow(gapindex_sizecomp), 
             ", crabpack ", nrow(crabpack_sizecomp), "), ", nrow(sizecomp_compare_diff), 
             " did not match. "))
# sizecomp_compare_diff
summary(sizecomp_compare_diff)
sizecomp_compare_diff_data <- sizecomp_compare |>
  dplyr::right_join(sizecomp_compare_diff |>
                      dplyr::select(survey_definition_id, year, species_code, length_mm, sex, noteable_diff))
# 
# sizecomp_compare_diff_data
# 
table(sizecomp_compare_diff_data$species_code, sizecomp_compare_diff_data$year, sizecomp_compare_diff_data$survey_definition_id)

write.csv(x = sizecomp_compare_diff_data, file = here::here("data/sizecomp_compare_diff_data.csv"))

```

Let's dive in to an example

Is the difference here just rounding error with the total area unit conversion?

```{r}
sizecomp_compare  |> 
  dplyr::rename_all(tolower) |>
  dplyr::filter(year == 2017 & species_code == 68560 & survey_definition_id  == 143) |> 
  dplyr::full_join(
    crab_specimen_preinclude |> 
      dplyr::rename_all(tolower) |> 
      dplyr::filter(year == 2017 & species_code == 68560 & region == 'NBS') |>
      dplyr::select(hauljoin, species_code, sex, length_mm, weight_g, haul_type, station_id) |> 
      dplyr::left_join(crabpack_cpue |> dplyr::select(-weight_kg, -count))
  )
#  |> 
# dplyr::mutate(
#   AREA_ID = ifelse(station_id == "AA-19", 71, 81), 
#   AREA_NM2 = ifelse(AREA_ID == 71, 57744 * x * 3.4299, # nmi2 to km2
#                     57744 * y * 3.4299, # nmi2 to km2
#                     ), # being used by crabpack
#   AREA_KM2 = ifelse(AREA_ID == 71, 81255.025, 38351.89), # from GAP_PRODUCTS
#   population_count.ytest = population_count.y/AREA_KM2
# )


```


