---
title: Access data via Oracle and R
number-sections: true
number-depth: 3
---

```{r setup, include=FALSE}
#| file: functions.R
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE, eval = TRUE, comment = FALSE)
```

## Access data via Oracle (AFSC only) {.unnumbered}

AFSC `Oracle` users can access the database via `SQL developer` to view and pull the production data directly from the `GAP_PRODUCTS` `Oracle` schema. The user can also use `SQL developer` to view and pull the GAP Products data directly from the `GAP_PRODUCTS` `Oracle` schema. 

### Connect to Oracle from R

Many users will want to access the data from `Oracle` using `R`. The user will need to install the `RODBC` `R` package and ask OFIS (IT) connect `R` to `Oracle`. Then, use the following code in `R` to establish a connection from `R` to `Oracle`: 

Here, the user can establish the oracle connection by entering their username and password in the `channel <- gapindex::oracle_connect()` function. Never save usernames or passwords in scripts that may be intentionally or unintentionally shared with others. If no username and password is entered in the function, pop-ups will appear on the screen asking for the username and password. 

After you connect to VPN, you'll be able to log into Oracle. 

```{r oracle-connect-2, echo = TRUE, eval = FALSE}
library(RODBC)
channel <- gapindex::get_connected()
```

## Data SQL Query Examples: {.unnumbered}

```{r load packages, echo = TRUE, eval = TRUE}
library(gapindex)
library(RODBC)
library(flextable)
library(ggplot2)
library(magrittr)
library(dplyr)
```

### Ex. Select all data from tables

You can download all of the tables locally using a variation of the code below. Once connected, pull and save the tables of interest into the `R` environment. 

```{r}
#| label: test-0
#| eval: false
#| message: false
#| warning: false
#| echo: true

locations <- c(
  "GAP_PRODUCTS.AKFIN_AGECOMP", 
  "GAP_PRODUCTS.AKFIN_AREA", 
  "GAP_PRODUCTS.AKFIN_BIOMASS", 
  "GAP_PRODUCTS.AKFIN_CATCH", 
  "GAP_PRODUCTS.AKFIN_CPUE", 
  "GAP_PRODUCTS.AKFIN_CRUISE", 
  "GAP_PRODUCTS.AKFIN_HAUL", 
  "GAP_PRODUCTS.AKFIN_LENGTH", 
  "GAP_PRODUCTS.AKFIN_METADATA_COLUMN", 
  "GAP_PRODUCTS.AKFIN_SIZECOMP", 
  "GAP_PRODUCTS.AKFIN_SPECIMEN", 
  "GAP_PRODUCTS.AKFIN_STRATUM_GROUPS", 
  "GAP_PRODUCTS.AKFIN_SURVEY_DESIGN", 
  "GAP_PRODUCTS.AKFIN_TAXONOMIC_CLASSIFICATION"
)

for (i in 1:length(locations)) {
  print(locations[i])
  a <- RODBC::sqlQuery(channel, paste0("SELECT * FROM ", locations[i]))
  write.csv(x = a, file = here::here("data", paste0(locations[i], ".csv")))
}
```

```{r}
#| label: test-11
#| message: false
#| warning: false
#| echo: true
#| eval: false
#| tbl-cap: "Pull AI rockfish with {dbplyr}. "

library(odbc)
library(RODBC)
library(dbplyr)

my_spp_codes <- c(
  30010, #	Sebastolobus sp.	thornyhead unid.
  30020, #	Sebastolobus alascanus	shortspine thornyhead
  30025, #	Sebastolobus macrochir	broadfin thornyhead
  30330, #	Sebastes melanops	black rockfish
  30430, #	Sebastes proriger	redstripe rockfish
  30470, #	Sebastes ruberrimus	yelloweye rockfish
  30475, #	Sebastes babcocki	redbanded rockfish
  30535, #	Sebastes variegatus	harlequin rockfish
  30560, # Sebastes zacentrus	sharpchin rockfish
  30600, # Sebastes reedi	yellowmouth rockfish
  30030, # Sebastolobus altivelis	longspine thornyhead
  30040, # Sebastes sp.	rockfish unid.
  30100, # Sebastes brevispinis	silvergray rockfish
  30150, # NA	dusky and dark rockfishes unid.
  30152, # Sebastes variabilis	dusky rockfish
  30170, # Sebastes crameri	darkblotched rockfish
  30270) # Sebastes helvomaculatus	rosethorn rockfish

a <- dplyr::tbl(channel, dplyr::sql('gap_products.akfin_biomass')) %>% 
  dplyr::rename_all(tolower) %>% 
  dplyr::select(survey_definition_id, area_id, species_code, year, biomass_mt, biomass_var) %>% 
  dplyr::filter(species_code %in% my_spp_codes & 
                  area_id %in% 99904 & 
                  year >= 1991) %>% 
  dplyr::collect() 

flextable::flextable(head(a)) %>% 
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra()
```


### Ex. CPUE for all EBS and NBS stations with associated haul, cruise, and species information.

```{r}
#| label: test-7
#| message: false
#| warning: false
#| echo: true
#| tbl-cap: "CPUE for all EBS and NBS stations with associated haul, cruise, and species information. "

a <- RODBC::sqlQuery(channel = channel, # NOT RACEBASE.HAUL
                     query = paste0(
                       "
-- Select columns for output data
SELECT
cr.CRUISEJOIN,
cr.CRUISE,
cr.YEAR,
cr.SURVEY_DEFINITION_ID,
cr.SURVEY_NAME,
cr.VESSEL_ID,
cr.VESSEL_NAME,
cp.HAULJOIN,
cp.SPECIES_CODE,
tt.SPECIES_NAME,
tt.COMMON_NAME,
cp.WEIGHT_KG,
cp.COUNT,
cp.AREA_SWEPT_KM2,
cp.CPUE_KGKM2,
cp.CPUE_NOKM2,
hh.HAUL,
hh.STATION

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_HAUL hh
LEFT JOIN GAP_PRODUCTS.AKFIN_CRUISE cr
ON hh.CRUISEJOIN = cr.CRUISEJOIN
LEFT JOIN GAP_PRODUCTS.AKFIN_CPUE cp
ON hh.HAULJOIN = cp.HAULJOIN
LEFT JOIN GAP_PRODUCTS.TAXONOMIC_CLASSIFICATION tt
ON cp.SPECIES_CODE = tt.SPECIES_CODE

-- Filter for EBS and NBS observations
WHERE SURVEY_DEFINITION_ID IN (143, 98) -- 143 NBS, 98 EBS
AND tt.SURVEY_SPECIES = 1

-- Only return the first 3 rows because otherwise this would be a huge table!
FETCH FIRST 3 ROWS ONLY;")) 

flextable::flextable(head(a[,2:8])) %>% 
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra()
```

### Ex. CPUE for all stations contained in the INPFC Shumagin region (AREA_ID = 919) for Pacific cod. 

```{r}
#| label: test-8
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel,
                       query =
                         "
-- Select columns for output data
SELECT 
HAULJOIN, 
SPECIES_CODE, 
STRATUM, 
LATITUDE_DD_START, 
LONGITUDE_DD_START,
CPUE_KGKM2, 
GEAR_TEMPERATURE_C

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_CPUE cpue
LEFT JOIN GAP_PRODUCTS.AKFIN_HAUL haul
USING (HAULJOIN) 

-- Filter for P. Cod observations
WHERE SPECIES_CODE IN (21720)

-- Select all stratum within the area_id 919 (INPFC Shumagin region)
AND haul.STRATUM IN
(
SELECT 
STRATUM
FROM GAP_PRODUCTS.AKFIN_STRATUM_GROUPS 
WHERE AREA_ID = 919
);")
```

```{r}
#| label: test-8-tbl
#| tbl-cap: "CPUE for all stations contained in the Shumagin region (AREA_ID = 919). "
#| message: false
#| warning: false
#| echo: true

dat <- dat %>% 
  dplyr::select(HAULJOIN, STRATUM, SPECIES_CODE, LATITUDE_DD_START, LONGITUDE_DD_START, CPUE_KGKM2, GEAR_TEMPERATURE_C) %>% 
  dplyr::mutate(SPECIES_CODE = as.character(SPECIES_CODE), 
                STRATUM = as.character(STRATUM)) %>% 
  dplyr::arrange(SPECIES_CODE)

flextable::flextable(head(dat)) %>%   
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra()
```

### Ex. EBS Pacific Ocean perch CPUE and [`akgfmaps`](https://github.com/afsc-gap-products/akgfmaps) map 

Pacific Ocean perch catch-per-unit-effort estimates for EBS in 2021 from `GAP_PRODUCTS.AKFIN_CPUE` and map constructed using [`akgfmaps`](https://github.com/afsc-gap-products/akgfmaps). Here, we'll use AKFIN HAUL and CRUISES data also included in this repo, for convenience, though they are very similar to their `RACEBASE` analogs. 

```{r}
#| label: test-6
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel, 
                       query = 
                         "
-- Select columns for output data
SELECT 
(cp.CPUE_KGKM2/100) CPUE_KGHA, -- akgfmaps is expecting hectares, but can take any units
hh.LATITUDE_DD_START LATITUDE,
hh.LONGITUDE_DD_START LONGITUDE

-- Use HAUL data to obtain LATITUDE & LONGITUDE and connect to cruisejoin
FROM GAP_PRODUCTS.AKFIN_CPUE cp
LEFT JOIN GAP_PRODUCTS.AKFIN_HAUL hh
ON cp.HAULJOIN = hh.HAULJOIN

-- Use CRUISES data to obtain YEAR and SURVEY_DEFINITION_ID
LEFT JOIN GAP_PRODUCTS.AKFIN_CRUISE cc
ON hh.CRUISEJOIN = cc.CRUISEJOIN

-- Filter data
WHERE cp.SPECIES_CODE = 30060 
AND cc.SURVEY_DEFINITION_ID = 98 
AND cc.YEAR = 2021;")
```

```{r}
#| label: test-6-tbl
#| tbl-cap: "EBS Pacific Ocean perch CPUE and [`akgfmaps`](https://github.com/afsc-gap-products/akgfmaps) map. "
#| message: false
#| warning: false
#| echo: true

dat %>% 
  dplyr::arrange(desc(CPUE_KGHA)) %>% 
  head() %>% 
  flextable::flextable() %>%  
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra()
```


```{r}
#| label: test-6-fig
#| fig-cap: "EBS Pacific Ocean perch CPUE and [`akgfmaps`](https://github.com/afsc-gap-products/akgfmaps) map. "
#| message: false
#| echo: true
#| warning: false

# devtools::install_github("afsc-gap-products/akgfmaps", build_vignettes = TRUE)
library(akgfmaps)

figure <- akgfmaps::make_idw_map(
  x = dat, # Pass data as a data frame
  region = "bs.south", # Predefined EBS area
  set.breaks = "jenks", # Gets Jenks breaks from classint::classIntervals()
  in.crs = "+proj=longlat", # Set input coordinate reference system
  out.crs = "EPSG:3338", # Set output coordinate reference system
  grid.cell = c(20000, 20000), # 20x20km grid
  key.title = "Pacific Ocean perch") # Include in the legend title

figure$plot
```

### Ex. GOA Pacific Ocean perch biomass and abundance

Biomass and abundance for Pacific Ocean perch from 1990 – 2023 for the western/central/eastern GOA management areas as well as for the entire region. 

```{r}
#| label: test-1
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel, 
                       query = 
                         "
-- Manipulate data to join to
WITH FILTERED_STRATA AS (
SELECT AREA_ID, DESCRIPTION FROM GAP_PRODUCTS.AKFIN_AREA
WHERE AREA_TYPE in ('REGULATORY_AREA', 'REGION') 
AND SURVEY_DEFINITION_ID = 47)

-- Select columns for output data
SELECT 
BIOMASS_MT,
POPULATION_COUNT, 
YEAR, 
DESCRIPTION

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_BIOMASS BIOMASS
JOIN FILTERED_STRATA STRATA 
ON STRATA.AREA_ID = BIOMASS.AREA_ID

-- Filter data results
WHERE BIOMASS.SPECIES_CODE = 30060")
```

```{r}
#| label: test-1-tbl
#| tbl-cap: "GOA Pacific Ocean perch biomass and abundance. "
#| message: false
#| warning: false
#| echo: true

dat0 <- dat %>% 
  janitor::clean_names() %>% 
  dplyr::select(biomass_mt, population_count, year, area = description) %>%
  pivot_longer(cols = c("biomass_mt", "population_count"), 
               names_to = "var", 
               values_to = "val") %>% 
  dplyr::mutate(
    val = ifelse(var == "biomass_mt", val/1e6, val/1e9), 
    var = ifelse(var == "biomass_mt", "Biomass (Mmt)", "Population (B)"), 
    area = gsub(x = area, pattern = " - ", replacement = "\n"), 
    area = gsub(x = area, pattern = ": ", replacement = "\n"), 
    type = sapply(X = strsplit(x = area, split = "\n", fixed = TRUE), `[[`, 2))  %>% 
  dplyr::arrange(type) %>% 
  dplyr::mutate(
    area = factor(area, levels = unique(area), labels = unique(area), ordered = TRUE))

flextable::flextable(head(dat)) %>% 
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra() %>%
  flextable::colformat_num(x = ., j = "YEAR", big.mark = "")
```

```{r}
#| label: test-1-plot
#| fig-cap: "GOA Pacific Ocean perch biomass and abundance. "
#| message: false
#| warning: false
#| echo: true

# install.packages("scales")
library(scales)
figure <- ggplot2::ggplot(
  dat = dat0, 
  mapping = aes(x = year, y = val, color = type)) +
  ggplot2::geom_point(size = 3) + 
  ggplot2::facet_grid(cols = vars(area), rows = vars(var), scales = "free_y") + 
  ggplot2::scale_x_continuous(name = "Year", n.breaks = 3) +
  ggplot2::scale_y_continuous(name = "Estimate", labels = comma) +
  ggplot2::labs(title = 'GOA Pacific Ocean perch biomass and abundance 1990 – 2023')  + 
  ggplot2::guides(color=guide_legend(title = "Region Type"))+
  ggplot2::scale_color_grey() +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.direction = "horizontal", 
                 legend.position = "bottom")

figure
```

### Ex. AI rock sole size compositions and ridge plot

Northern and Southern rock sole size composition data from 1991 – 2022 for the Aleutian Islands, with Ridge plot from [`ggridges`](https://cran.r-project.org/web/packages/ggridges/vignettes/introduction.html).

```{r}
#| label: test-2
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel, 
                       query = "
SELECT 
YEAR,
LENGTH_MM / 10 AS LENGTH_CM, 
SUM(POPULATION_COUNT) AS POPULATION_COUNT

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_SIZECOMP 

-- 99904 is the AREA_ID that codes for the whole AI survey region
WHERE AREA_ID = 99904
-- including northern rock sole, southern rock sole, and rock sole unid.
AND SPECIES_CODE IN (10260, 10261, 10262)
-- remove the -9 LENGTH_MM code
AND LENGTH_MM > 0
-- sum over species_codes and sexes
GROUP BY (YEAR, LENGTH_MM)")
```

```{r}
#| label: test-2-tbl
#| tbl-cap: "AI Rock sole size compositions and ridge plot. "
#| message: false
#| warning: false
#| echo: true

dat0 <- dat %>% 
  janitor::clean_names() %>% 
  head() %>% 
  flextable::flextable() %>% 
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra() %>%
  flextable::colformat_num(x = ., j = "year", big.mark = "")
dat0
```


```{r}
#| label: test-2-plot
#| fig-cap: "AI Rock sole size compositions and ridge plot. "
#| message: false
#| warning: false
#| echo: true

# install.packages("ggridges")
library(ggridges)
figure <- ggplot(dat, 
                 mapping = aes(x = LENGTH_CM, 
                               y = YEAR, 
                               height = POPULATION_COUNT, 
                               group = YEAR)) +
  ggridges::geom_density_ridges(stat = "identity", scale = 1) +
  ggplot2::ylab(label = "Year") +
  ggplot2::scale_x_continuous(name = "Length (cm)") +
  ggplot2::labs(title = paste0('Aleutian Islands Rock sole Size Compositions'), 
                subtitle = paste0(min(dat$YEAR), ' – ', max(dat$YEAR))) +
  ggplot2::theme_bw()

figure
```

### Ex. 2023 EBS Walleye Pollock Age Compositions and Age Pyramid

Walleye pollock age composition for the EBS standard + NW Area from 2023, with age pyramid plot. 

```{r}
#| label: test-3
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel, 
                       query = "
-- Manipulate data to join to
WITH FILTERED_STRATA AS (
SELECT 
AREA_ID, 
DESCRIPTION 
FROM GAP_PRODUCTS.AKFIN_AREA
WHERE AREA_TYPE = 'REGION' AND 
SURVEY_DEFINITION_ID = 98)

-- Select columns for output data
SELECT 
AGECOMP.AGE, 
AGECOMP.POPULATION_COUNT, 
AGECOMP.SEX

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_AGECOMP AGECOMP
JOIN FILTERED_STRATA STRATA 
ON STRATA.AREA_ID = AGECOMP.AREA_ID

-- Filter data results
WHERE SPECIES_CODE = 21740
AND YEAR = 2023
AND AGE >= 0")
```

```{r}
#| label: test-3-tbl
#| tbl-cap: "EBS Walleye Pollock Age Compositions and Age Pyramid. "
#| message: false
#| warning: false
#| echo: true

dat0 <- dat %>% 
  janitor::clean_names() %>% 
  dplyr::filter(sex %in% c(1,2)) %>%
  dplyr::mutate(
    sex = ifelse(sex == 1, "M", "F"),
    population_count = # change male population to negative
      ifelse(sex=="M", population_count*(-1), population_count*1)/1e9) 

flextable::flextable(head(dat)) %>% 
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra()
```


```{r}
#| label: test-3-plot
#| fig-cap: "2023 EBS Walleye Pollock Age Compositions and Age Pyramid. "
#| message: false
#| warning: false
#| echo: true

figure <- ggplot2::ggplot(
  data = dat0, 
  mapping = 
    aes(x = age,
        y = population_count, 
        fill = sex)) +
  ggplot2::scale_fill_grey() +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::coord_flip() +
  ggplot2::scale_x_continuous(name = "Age") +
  ggplot2::scale_y_continuous(name = "Population (billions)", labels = abs) +
  ggplot2::ggtitle(label = "2023 EBS Walleye Pollock Age Compositions")  + 
  ggplot2::guides(fill = guide_legend(title = "Sex"))+
  ggplot2::theme_bw()

figure
```

### Ex. NBS Pacific cod biomass and abundance

Pacific cod biomass and abundance data for the NBS by stratum. 

```{r}
#| label: test-4
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel, 
                       query = 
                         "
-- Manipulate data to join to
WITH FILTERED_STRATA AS (
SELECT 
AREA_ID, 
AREA_NAME, 
DESCRIPTION 
FROM GAP_PRODUCTS.AKFIN_AREA
WHERE AREA_TYPE in ('STRATUM') AND 
SURVEY_DEFINITION_ID = 143) 

-- Select columns for output data
SELECT 
BIOMASS.BIOMASS_MT, 
BIOMASS.POPULATION_COUNT, 
BIOMASS.YEAR, 
STRATA.AREA_NAME

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_BIOMASS BIOMASS 
JOIN FILTERED_STRATA STRATA 
ON STRATA.AREA_ID = BIOMASS.AREA_ID

-- Filter data results
WHERE BIOMASS.SURVEY_DEFINITION_ID IN 143 
AND BIOMASS.SPECIES_CODE = 21720")
```

```{r}
#| label: test-4-tbl
#| tbl-cap: "NBS Pacific cod biomass and abundance. "
#| message: false
#| warning: false
#| echo: true

dat0 <- dat %>% 
  janitor::clean_names() %>% 
  dplyr::select(biomass_mt, population_count, year, area = area_name) %>%
  pivot_longer(cols = c("biomass_mt", "population_count"), 
               names_to = "var", 
               values_to = "val") %>% 
  dplyr::mutate(
    val = ifelse(var == "biomass_mt", val/1e6, val/1e9), 
    var = ifelse(var == "biomass_mt", "Biomass (Mmt)", "Population (B)"), 
    area = factor(area, levels = unique(area), labels = unique(area), ordered = TRUE))
flextable::flextable(head(dat)) %>% 
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra() %>%
  flextable::colformat_num(x = ., j = "YEAR", big.mark = "")
```

```{r}
#| label: test-4-fig
#| fig-cap: "NBS Pacific cod biomass and abundance. "
#| message: false
#| warning: false
#| echo: true

figure <- ggplot2::ggplot(
  dat = dat0, 
  mapping = aes(y = val, x = year, fill = area))  + 
  ggplot2::geom_bar(position="stack", stat="identity") +  
  ggplot2::facet_grid(rows = vars(var), scales = "free_y") +
  ggplot2::scale_y_continuous(name = "Estimate", labels = comma) +
  ggplot2::scale_x_continuous(name = "Year", breaks = unique(dat0$year)) +
  ggplot2::labs(title = 'NBS Pacific cod biomass and abundance by stratum')  + 
  ggplot2::guides(fill=guide_legend(title = "Region Type"))+
  ggplot2::scale_fill_grey() +
  ggplot2::theme_bw() +
  ggplot2::theme(legend.direction = "horizontal", 
                 legend.position = "bottom")

figure
```

### Ex. GOA Pacific Ocean perch biomass and line plot

Pacific Ocean perch biomass totals for GOA between 1984-2021 from `GAP_PRODUCTS.AKFIN_BIOMASS`

```{r}
#| label: test-5
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel,
                       query = "
-- Select columns for output data
SELECT
SURVEY_DEFINITION_ID,
BIOMASS_MT / 1000 AS BIOMASS_KMT,
(BIOMASS_MT - 2 * SQRT(BIOMASS_VAR)) / 1000 AS BIOMASS_KCI_DW,
(BIOMASS_MT + 2 * SQRT(BIOMASS_VAR)) / 1000 AS BIOMASS_KCI_UP,
YEAR

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_BIOMASS

-- Filter data results
WHERE SPECIES_CODE = 30060
AND SURVEY_DEFINITION_ID = 47
AND AREA_ID = 99903
AND YEAR BETWEEN 1990 AND 2023" ) %>% 
  janitor::clean_names()
```

```{r}
#| label: test-5-tbl
#| tbl-cap: "GOA Pacific Ocean perch biomass and line plot. "
#| message: false
#| warning: false
#| echo: true

flextable::flextable(head(dat)) %>%
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra() %>%
  flextable::colformat_num(x = ., j = "year", big.mark = "")
```

```{r}
#| label: test-5-fig
#| fig-cap: "GOA Pacific Ocean perch biomass and line plot. "
#| message: false
#| warning: false
#| echo: true

a_mean <- dat %>% 
  dplyr::group_by(survey_definition_id) %>% 
  dplyr::summarise(biomass_kmt = mean(biomass_kmt, na.rm = TRUE), 
                   minyr = min(year, na.rm = TRUE), 
                   maxyr = max(year, na.rm = TRUE)) 

figure <-
  ggplot(data = dat, 
         mapping = aes(x = year, 
                       y = biomass_kmt)) +
  ggplot2::geom_point(size = 2.5, color = "grey40") + 
  ggplot2::scale_x_continuous(
    name = "Year", 
    labels = scales::label_number(
      accuracy = 1, 
      big.mark = ""))   +
  ggplot2::scale_y_continuous(
    name = "Biomass (Kmt)", 
    labels = comma) +
  ggplot2::geom_segment(
    data = a_mean,
    mapping = aes(x = minyr, 
                  xend = maxyr, 
                  y = biomass_kmt, 
                  yend = biomass_kmt),
    linetype = "dashed", 
    linewidth = 2) +
  ggplot2::geom_errorbar(
    mapping = aes(ymin = biomass_kci_dw, ymax = biomass_kci_up),
    position = position_dodge(.9),
    alpha = 0.5, width=.2) +
  ggplot2::ggtitle(
    label = "GOA Pacific Ocean Perch Biomass 1984-2021", 
    subtitle = paste0("Mean = ", 
                      formatC(x = a_mean$biomass_kmt, 
                              digits = 2, 
                              big.mark = ",", 
                              format = "f"), 
                      " Kmt")) +
  ggplot2::theme_bw()

figure
```

### Ex. 2022 AI Atka mackerel age specimen summary

#### All ages determined:

```{r}
#| label: test-10a
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel,
                       query = "
-- Select columns for output data
SELECT SURVEY_DEFINITION_ID, YEAR, SPECIES_CODE, AGE

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_SPECIMEN
JOIN (SELECT HAULJOIN, CRUISEJOIN FROM GAP_PRODUCTS.AKFIN_HAUL)
USING (HAULJOIN)
JOIN (SELECT CRUISEJOIN, YEAR, SURVEY_DEFINITION_ID FROM GAP_PRODUCTS.AKFIN_CRUISE)
USING (CRUISEJOIN)

-- Filter data results
WHERE GAP_PRODUCTS.AKFIN_SPECIMEN.SPECIMEN_SAMPLE_TYPE = 1
AND SPECIES_CODE = 21921
AND YEAR = 2022
AND SURVEY_DEFINITION_ID = 52") %>% 
  janitor::clean_names()
```

```{r}
#| label: test-10a-tbl
#| tbl-cap: "2022 AI Atka mackerel age specimen summary: all ages determined. "
#| message: false
#| warning: false
#| echo: true

flextable::flextable(head(dat) %>% 
                       dplyr::arrange(age)) %>%
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra() %>%
  flextable::colformat_num(x = ., j = c("year", "species_code"), big.mark = "")
```

#### How many of each age was found:

```{r}
#| label: test-10b
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel,
                       query = "
-- Select columns for output data
SELECT SURVEY_DEFINITION_ID, YEAR, SPECIES_CODE, AGE, 
COUNT(AGE) AS COUNTAGE

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_SPECIMEN
JOIN (SELECT HAULJOIN, CRUISEJOIN FROM GAP_PRODUCTS.AKFIN_HAUL)
USING (HAULJOIN)
JOIN (SELECT CRUISEJOIN, YEAR, SURVEY_DEFINITION_ID FROM GAP_PRODUCTS.AKFIN_CRUISE)
USING (CRUISEJOIN)

-- Filter data results
WHERE GAP_PRODUCTS.AKFIN_SPECIMEN.SPECIMEN_SAMPLE_TYPE = 1
AND SPECIES_CODE = 21921
AND YEAR = 2022
AND SURVEY_DEFINITION_ID = 52
GROUP BY (YEAR, SURVEY_DEFINITION_ID, SPECIES_CODE, AGE)") %>% 
  janitor::clean_names()
```

```{r}
#| label: test-10b-tbl
#| tbl-cap: "Ex.: 2022 AI Atka mackerel age specimen summary: how many of each age were determined. "
#| message: false
#| warning: false
#| echo: true

flextable::flextable(head(dat) %>% 
                       dplyr::arrange(age)) %>%
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra() %>%
  flextable::colformat_num(x = ., j = c("year", "species_code"), big.mark = "")
```

#### How many otoliths were aged: 

Using `SQL`

```{r}
#| label: test-10c1
#| message: false
#| warning: false
#| echo: true

dat <- RODBC::sqlQuery(channel = channel,
                       query = "
-- Select columns for output data
SELECT SURVEY_DEFINITION_ID, YEAR, SPECIES_CODE, 
COUNT(AGE) AS COUNTAGE

-- Identify what tables to pull data from
FROM GAP_PRODUCTS.AKFIN_SPECIMEN
JOIN (SELECT HAULJOIN, CRUISEJOIN FROM GAP_PRODUCTS.AKFIN_HAUL)
USING (HAULJOIN)
JOIN (SELECT CRUISEJOIN, YEAR, SURVEY_DEFINITION_ID FROM GAP_PRODUCTS.AKFIN_CRUISE)
USING (CRUISEJOIN)

-- Filter data results
WHERE GAP_PRODUCTS.AKFIN_SPECIMEN.SPECIMEN_SAMPLE_TYPE = 1
AND SPECIES_CODE = 21921
AND YEAR = 2022
AND SURVEY_DEFINITION_ID = 52
GROUP BY (YEAR, SURVEY_DEFINITION_ID, SPECIES_CODE)") %>% 
  janitor::clean_names()
```

Using `dbplyr`:

```{r}
#| label: test-10c2
#| message: false
#| warning: false
#| echo: true
#| eval: false

library(odbc)
library(keyring)
library(dplyr)
library(dbplyr)

channel <- DBI::dbConnect(odbc::odbc(), "akfin", uid = keyring::key_list("akfin")$username,
                        pwd = keyring::key_get("akfin", keyring::key_list("akfin")$username))

dat <- dplyr::tbl(src = channel, dplyr::sql('gap_products.akfin_specimen')) %>% 
    dplyr::rename_all(tolower) %>% 
    dplyr::select(hauljoin, specimen = specimen_id, species_code, length = length_mm, 
                  weight = weight_g, age, sex, age_method = age_determination_method) %>% 
    dplyr::left_join(dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_haul')) %>%
                       dplyr::rename_all(tolower) %>% 
                       dplyr::select(cruisejoin, hauljoin, haul, date_collected = date_time_start, 
                                     latitude = latitude_dd_start, longitude = longitude_dd_start),
                     by = join_by(hauljoin)) %>% 
    dplyr::left_join(dplyr::tbl(akfin, dplyr::sql('gap_products.akfin_cruise')) %>%
                       dplyr::rename_all(tolower) %>% 
                       dplyr::select(cruisejoin, year, vessel = vessel_id, survey_definition_id),
                     by = join_by(cruisejoin)) %>% 
    dplyr::filter(year == YEAR &
             survey_definition_id == 52 & 
             species_code %in% spp_codes &
             !is.na(age)) %>% 
    dplyr::collect()
```

Both scripts will produce this table: 

```{r}
#| label: test-10c1-tbl
#| tbl-cap: "2022 AI Atka mackerel age specimen summary: how many otoliths were aged. This quiery was created using SQL. "
#| message: false
#| warning: false
#| echo: true

flextable::flextable(head(dat)) %>%
  flextable::fit_to_width(max_width = 6) %>% 
  flextable::theme_zebra() %>%
  flextable::colformat_num(x = ., j = c("year", "species_code"), big.mark = "")
```



```{r rodbc-close}
RODBC::odbcCloseAll()
```
