---
title: Public Data (FOSS)
---

```{r setup, include=FALSE}
#| file: functions.R
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE, eval = TRUE, comment = FALSE)
```

The final, validated survey data are publicly accessible soon after surveys are completed on the [Fisheries One Stop Shop (FOSS) platform](https://www.fisheries.noaa.gov/foss/). This data includes catch, haul, and environmental data collected at each station. On the [FOSS data platform](https://www.fisheries.noaa.gov/foss/), users can interactively select, view, and download data. Descriptive documentation and user-examples are available on the metadata page.

This data contains all of the catch, environmental, and haul data from the fisheries-independent Groundfish and Shellfish Assessment Program surveys in the Bering Sea, Aleutian Islands, and Gulf of Alaska. This data is sought after by the general public, private entities, and NOAA partners alike, including tribal organizations, K-12 classrooms, academic institutions, for-profit groups, and non-profit groups. This data is compiled and approved once a year after each summer survey season and is available for open access. 

# Collaborators and data users

Below are a few packages and products currently using this data. If you have developed a product, performed an analysis, or exhibited this data in any way, reach out so we can showcase your hard work. 

```{r citing-us, echo = FALSE}
str0 <- paste0(" - **[NOAA Fisheries Distribution Mapping and Analysis Portal](https://apps-st.fisheries.noaa.gov/dismap)**; *[NOAA Fisheries Office of Science and Technology](https://www.fisheries.noaa.gov/contact/office-science-and-technology)*


 - **[Pull data with python](https://pyafscgap.org/) and explore the [in-browser visualization tool](https://app.pyafscgap.org/'). Reference their [example Python notebook](https://mybinder.org/v2/gh/SchmidtDSE/afscgap/main?urlpath=/tree/index.ipynb)**; *[The Eric and Wendy Schmidt Center for Data Science and the Environment at UC Berkeley](https://dse.berkeley.edu/), including sam.pottinger@berkeley.edu, ccmartinez@berkeley.edu, gzarpellon@berkeley.edu, and kkoy@berkeley.edu.* ")

#  # DisMAP
# df <- data.frame(    
#     url = c("https://www.noaa.gov/sites/default/files/2022-03/noaa_emblem_logo-2022.png"), 
#     Description = c("<a href = 'https://apps-st.fisheries.noaa.gov/dismap'>NOAA Fisheries Distribution Mapping and Analysis Portal</a>"), 
#     Contacts = c("<a href = 'https://www.fisheries.noaa.gov/contact/office-science-and-technology'>NOAA Fisheries Office of Science and Technology</a>"))
# 
#  # DSE
# df <- df |> 
#   dplyr::add_row(
#     url = c("https://dse.berkeley.edu/sites/default/files/styles/openberkeley_image_full/public/general/dse_logostack.png"), 
#                  Description = c("<a href = 'https://pyafscgap.org/'>Pull data with python</a> and explore the <a href = 'https://app.pyafscgap.org/'>in-browser visualization tool</a>. Reference their <a href = 'https://mybinder.org/v2/gh/SchmidtDSE/afscgap/main?urlpath=/tree/index.ipynb'>example Python notebook</a>"), 
#                  Contacts = c("<a href = 'https://dse.berkeley.edu/'>The Eric and Wendy Schmidt Center for Data Science and the Environment at UC Berkeley</a>; <a href = 'mailto: sam.pottinger@berkeley.edu'>Sam Pottinger</a>, <a href = 'mailto: ccmartinez@berkeley.edu'>Ciera Martinez</a>, <a href = 'mailto: gzarpellon@berkeley.edu'>Giulia Zarpellon</a>, and <a href = 'mailto: kkoy@berkeley.edu'>Kevin Koy</a>"))
# 
# library(gt)
# library(gtExtras)
# library(purrr)
#   
#   df <- df |>
#   dplyr::select(url, 
#                 Description, Contacts) |>
#   dplyr::mutate(Contacts = map(Contacts, gt::html)) |>
#   dplyr::mutate(Description = map(Description, gt::html)) |>
#   gt::gt() |>
#   gtExtras::gt_img_rows(columns = url, img_source = "web") |>
#   gt::tab_options(data_row.padding = px(1)) |>
#   gt::cols_label(url = "     ") |>
#   cols_align(align = c("left"), columns = everything())
# df

```

`r str0`

## Access Constraints

**User Constraints:** Users must read and fully comprehend the metadata prior to use. Data should not be used beyond the limits of the source scale. Acknowledgment of AFSC Groundfish Assessment Program, as the source from which these data were obtained, in any publications and/or other representations of these data, is suggested.

**General questions and more specific data requests** can be sent to <nmfs.afsc.gap.metadata@noaa.gov> or submitted as an [issue on our GitHub
Organization](https://github.com/afsc-gap-products/data-requests). The version of this data used for stock assessments can be found through the Alaska Fisheries Information Network (AKFIN). For questions about the eastern Bering Sea surveys, contact Duane Stevenson (<Duane.Stevenson@noaa.gov>). For questions about the Gulf of Alaska or Aleutian Islands surveys, contact Ned Laman (<Ned.Laman@noaa.gov>). For questions specifically about crab data in any region, contact Mike Litzow (<Mike.Litzow@noaa.gov>), the Shellfish Assessment Program lead.

For questions, comments, and concerns specifically about the [Fisheries One Stop Shop (FOSS)](https://www.fisheries.noaa.gov/foss) platform, please contact us using the Comments page on the [FOSS](https://www.fisheries.noaa.gov/foss) webpage.

## Cite this data

Use the below bibtext citation, as cited in our group's [citation repository](https://github.com/afsc-gap-products/citations/blob/main/cite/bibliography.bib) for citing the data created and maintained in this repository. Add "note = {Accessed: mm/dd/yyyy}" to append the day this data was accessed. 

```{r load_citation, echo = FALSE}
find_citation_for(bib_ref = "FOSSAFSCData")
```

```{r rodbc-close}
RODBC::odbcCloseAll()
```
