---
title: Access via Oracle and R (AFSC only)
number-sections: true
---

```{r setup, include=FALSE}
#| file: functions.R
```

If the user has access to the AFSC `Oracle` database, the user can use `SQL developer` to view and pull the FOSS public data directly from the `GAP_PRODUCTS` `Oracle` schema. 

### Connect to Oracle from R

Many users will want to access the data from `Oracle` using `R`. The user will need to install the `RODBC` `R` package and ask OFIS (IT) connect `R` to `Oracle`. Then, use the following code in `R` to establish a connection from `R` to `Oracle`: 

Here, the user can write in their username and password directly into the `RODBC` connect function. Never save usernames or passwords in scripts that may be intentionally or unintentionally shared with others. If no username and password is entered in the function, pop-ups will appear on the screen asking for the username and password. 

```{r oracle-connect-2, echo = TRUE, eval = FALSE}
library(gapindex)
channel <- gapindex::get_connected()
```


### Ex. Wholesale download data and join data in R

```{r}
#| label: test-1a
#| tbl-cap: "Wholesale download data and join data in R. "
#| message: false
#| warning: false
#| echo: true
#| eval: false

locations <- c(
  "GAP_PRODUCTS.FOSS_CATCH",
  "GAP_PRODUCTS.FOSS_HAUL",
  "GAP_PRODUCTS.FOSS_SPECIES"
)

print(Sys.Date())

error_loading <- c() # log if any tables are unable to download 
for (i in 1:length(locations)){
  print(locations[i])
  a <- RODBC::sqlQuery(channel, paste0("SELECT * FROM ", locations[i], "; "))
  if (is.null(nrow(a))) { # if an error in downloading has occurred
    error_loading <- c(error_loading, locations[i])
  } else { # if no error in downloading has occurred
    write.csv(x = a, 
              # change file name to be more computer file storage friendly
              here::here(paste0(tolower(gsub(
                pattern = '.', 
                replacement = "_", 
                x = locations[i], 
                fixed = TRUE)),
                ".csv")))
  }
}
error_loading
```

Join downloaded files

```{r}
#| label: test-1b
#| tbl-cap: "Wholesale download data and join data in R. "
#| message: false
#| warning: false
#| echo: true
#| eval: false

# Load data
library(dplyr)
library(here)
library(readr)
catch <- readr::read_csv(file = here::here("data/gap_products_foss_catch.csv"))[,-1] # remove "row number" column
haul <- readr::read_csv(file = here::here("data/gap_products_foss_haul.csv"))[,-1] # remove "row number" column
species <- readr::read_csv(file = here::here("data/gap_products_foss_species.csv"))[,-1] # remove "row number" column

# come up with full combination of what species should be listed for what hauls/surveys
# for zero-filled data, all species caught in a survey need to have zero or non-zero row entries for a haul
comb <- dplyr::full_join(
  x = dplyr::left_join(catch, haul, by = "HAULJOIN") %>%
    dplyr::select(SURVEY_DEFINITION_ID, SPECIES_CODE) %>%
    dplyr::distinct(),
  y = haul %>%
    dplyr::select(SURVEY_DEFINITION_ID, HAULJOIN) %>%
    dplyr::distinct(), 
  by = "SURVEY_DEFINITION_ID", 
  relationship = "many-to-many"
)

# Join data to make a full zero-filled CPUE dataset
dat <- comb %>% 
  # add species data to unique species by survey table
  dplyr::left_join(species, "SPECIES_CODE") %>% 
  # add catch data
  dplyr::full_join(catch, c("SPECIES_CODE", "HAULJOIN")) %>% 
  # add haul data
  dplyr::full_join(haul) %>% # , c("SURVEY_DEFINITION_ID", "HAULJOIN")
  # modify zero-filled rows
  dplyr::mutate(
    CPUE_KGKM2 = ifelse(is.null(CPUE_KGKM2), 0, CPUE_KGKM2),
    CPUE_KGHA = CPUE_KGKM2/100, # Hectares
    CPUE_NOKM2 = ifelse(is.null(CPUE_NOKM2), 0, CPUE_NOKM2),
    CPUE_NOHA = CPUE_NOKM2/100, # Hectares
    COUNT = ifelse(is.null(COUNT), 0, COUNT),
    WEIGHT_KG = ifelse(is.null(WEIGHT_KG), 0, WEIGHT_KG) ) 
```

### Ex. Join data using Oracle

To join these tables in Oracle, you may use a variant of the following code: 

```{sql test-4, eval = FALSE, echo = TRUE}

SELECT 
hh.YEAR,
hh.SRVY,                 
hh.SURVEY,
hh.SURVEY_DEFINITION_ID,
hh.SURVEY_NAME,
hh.CRUISE,
hh.CRUISEJOIN,           
hh.HAUL,
hh.HAULJOIN,
hh.STRATUM,
hh.STATION,
hh.VESSEL_ID,
hh.VESSEL_NAME,          
hh.DATE_TIME,
hh.LATITUDE_DD_START, 
hh.LONGITUDE_DD_START, 
hh.LATITUDE_DD_END,
hh.LONGITUDE_DD_END, 
hh.BOTTOM_TEMPERATURE_C,
hh.SURFACE_TEMPERATURE_C,
hh.DEPTH_M,
cc.SPECIES_CODE,
ss.ITIS,
ss.WORMS,
ss.COMMON_NAME,     
ss.SCIENTIFIC_NAME,
ss.ID_RANK,
CASE WHEN cc.CPUE_KGKM2 IS NULL THEN 0 ELSE cc.CPUE_KGKM2 END AS CPUE_KGKM2,
CASE WHEN cc.CPUE_NOKM2 IS NULL THEN 0 ELSE cc.CPUE_NOKM2 END AS CPUE_NOKM2,
CASE WHEN cc.COUNT IS NULL THEN 0 ELSE cc.COUNT END AS COUNT,
CASE WHEN cc.WEIGHT_KG IS NULL THEN 0 ELSE cc.WEIGHT_KG END AS WEIGHT_KG,
CASE WHEN cc.TAXON_CONFIDENCE IS NULL THEN NULL ELSE cc.TAXON_CONFIDENCE END AS TAXON_CONFIDENCE,
hh.AREA_SWEPT_KM2,       
hh.DISTANCE_FISHED_KM,
hh.DURATION_HR,          
hh.NET_WIDTH_M,
hh.NET_HEIGHT_M,
hh.PERFORMANCE 
FROM GAP_PRODUCTS.FOSS_SURVEY_SPECIES sv
FULL OUTER JOIN GAP_PRODUCTS.FOSS_SPECIES ss
ON sv.SPECIES_CODE = ss.SPECIES_CODE
FULL OUTER JOIN GAP_PRODUCTS.FOSS_HAUL hh
ON sv.SURVEY_DEFINITION_ID = hh.SURVEY_DEFINITION_ID
FULL OUTER JOIN GAP_PRODUCTS.FOSS_CATCH cc
ON sv.SPECIES_CODE = cc.SPECIES_CODE
AND hh.HAULJOIN = cc.HAULJOIN
```

### Ex. Subset data

Here, we are pulling EBS Pacific cod from 2010 - 2021: 

```{r test-2, echo = TRUE, eval = TRUE}
# Pull data
a <- RODBC::sqlQuery(
channel = channel, 
query = 
"SELECT * FROM GAP_PRODUCTS.FOSS_CATCH cc
JOIN GAP_PRODUCTS.FOSS_HAUL hh
ON cc.HAULJOIN = hh.HAULJOIN
WHERE SRVY = 'EBS' 
AND SPECIES_CODE = 21720 -- 'Pacific cod' 
AND YEAR >= 2010 
AND YEAR < 2021")

head(a)
```

### Ex. Find all species found in the eastern Bering Sea (EBS) survey in 2023

```{r test-3, echo = TRUE, eval = TRUE}
# Pull data
a <- RODBC::sqlQuery(
channel = channel, 
query = 
"SELECT DISTINCT 
ss.COMMON_NAME,
ss.SCIENTIFIC_NAME, 
ss.ID_RANK, 
ss.WORMS
FROM GAP_PRODUCTS.FOSS_CATCH cc -- get species codes
LEFT JOIN GAP_PRODUCTS.FOSS_SPECIES ss -- get species info
ON cc.SPECIES_CODE = ss.SPECIES_CODE
LEFT JOIN GAP_PRODUCTS.FOSS_HAUL hh -- filter by year and survey
ON cc.HAULJOIN = hh.HAULJOIN
WHERE hh.YEAR = 2023
AND hh.SURVEY_DEFINITION_ID = 98 -- EBS survey
ORDER BY COMMON_NAME")

head(a)
```

```{r rodbc-close}
RODBC::odbcCloseAll()
```

