---
title: Introduction
---

```{r setup, include=FALSE}
#| file: functions.R
```

## Our Objective

These data are part of the public domain, and thus need to be available to the public. As part of our commitment to open science and transparency, we additionally provide this interactive metadata guide. Please refer to our [Draft Data Changes Brief](https://docs.google.com/document/d/1ie0it6G_V_PrpO1fYe-731Fvubuahn2yi0ixBIpaYO8/edit?usp=sharing). Once finalized, this language will be included here. 

## User Resources

-   [GitHub repository](`r link_repo`).
-   [Access Tips and Documentation for All Production Data](`r link_repo_web `)
-   [Fisheries One Stop Shop (FOSS)](`r link_foss`)

-   [Groundfish Assessment Program Bottom Trawl Surveys](https://www.fisheries.noaa.gov/alaska/science-data/groundfish-assessment-program-bottom-trawl-surveys)
-   [AFSC’s Resource Assessment and Conservation Engineering Division](https://www.fisheries.noaa.gov/about/resource-assessment-and-conservation-engineering-division)
-   [Survey code books](`r link_code_books `)
-   [Publications and Data Reports](https://repository.library.noaa.gov/)
-   [Research Surveys conducted at AFSC](https://www.fisheries.noaa.gov/alaska/ecosystems/alaska-fish-research-surveys)

## Cite this data

Use the below [bibtext citations](`r paste0(link_repo, "/blob/main/CITATION.bib")`), as cited in our group's [citation repository](https://github.com/afsc-gap-products/citations/blob/main/cite/bibliography.bib) for citing the data created and maintained in this repo. Add "note = {Accessed: mm/dd/yyyy}" to append the day this data was accessed. Included here are AFSC RACE Groundfish and Shellfish Assessment Program's: 
 
 - Design-Based Production Data (internal) [@GAPProducts].  
 - AFSC RACE Groundfish Data for AKFIN [@GAPakfin]. 
 - Public Data hosted on the Fisheries One Stop Shop (FOSS) Data Platform [@FOSSAFSCData]. 

```{r load_citation, echo = FALSE}
cat(readLines(con = here::here("CITATION.bib")), sep = "\n") 
```

Or cite our latest data reports for survey-specific data and other findings:

<div id="refs"></div>
  
## Access Constraints

There are no legal restrictions on access to the data. They reside in public domain and can be freely distributed.

**User Constraints:** Users must read and fully comprehend the metadata prior to use. Data should not be used beyond the limits of the source scale. Acknowledgement of AFSC Groundfish Assessment Program, as the source from which these data were obtained, in any publications and/or other representations of these data, is suggested.

