---
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE, error = FALSE)
```

```{r setup1}
library(kableExtra)
library(magrittr)
library(flextable)

this_yr <- 2023
past_yr <- this_yr-1

# find_difference <- function(mismatch, str0, past_yr, this_yr) {
# aa <- mismatch %>% 
#   dplyr::select(dplyr::ends_with("_hist")) %>% 
#   names()
# aa <- gsub(pattern = "_hist", replacement = "", x = aa)
# 
# a$changes <- ""
# for (i in 1:nrow(a)) {
#   aaaa <- c()
#   for (ii in 1:length(aa)) {
#     aaa <- a[i,] %>% 
#       dplyr::select(dplyr::starts_with(aa[ii])) 
#     if (aaa[,1] %in% aaa[,2]) {
#       aaaa <- c(aaaa, aa[ii])
#     }
#   }
#   aaaa <- ifelse(aaa[,1] == aaa[,2], "", 
#                 paste0(aaaa, collapse = ", "))
#   
#   a$changes[i] <- aaaa
# }
# 
# str00 <- paste0("There were ", nrow(a), 
#                 " ",str0," estimates that changed from the ", 
#                 past_yr," and ", this_yr," datasets. ")
# 
# return(list("a" = a, 
#             "str00" = str00))
# }


find_difference <- function(mismatch, str0, past_yr, this_yr, col) {

mismatch0 <- mismatch

aa <- mismatch %>% 
  dplyr::select(dplyr::ends_with("_hist")) %>% 
  names()
aa <- gsub(pattern = "_hist", replacement = "", x = aa)

mismatch <- mismatch %>% 
  dplyr::select(dplyr::any_of(col), 
                dplyr::ends_with("_hist"), 
                dplyr::ends_with("_prod"))  %>% 
  dplyr::select(order(colnames(.))) %>% 
  dplyr::relocate(dplyr::any_of(col))

for (i in 1:length(aa)) {
  aaa <- names(mismatch)[which(grepl(pattern = aa[i], x = names(mismatch)))]
  mismatch$temp <- (mismatch[, aaa[1]] - mismatch[, aaa[2]])
  names(mismatch)[names(mismatch) == "temp"] <- paste0(aa[i], "_diff")
  
  mismatch$temp <- #mismatch %>% 
    dplyr::case_when(
      is.na(mismatch[,aaa[1]]) ~ paste0("missing from hist"), 
      is.na(mismatch[,aaa[2]]) ~ paste0("missing from prod"), 
      abs(mismatch[,paste0(aa[i], "_diff")])>0.1 ~ paste0("rounding error in ", aa[i]), 
      abs(mismatch[,paste0(aa[i], "_diff")])<=0.1 ~ paste0("actual change in ", aa[i]), 
      # ~ "FALSE"
    )
  
  # mismatch$temp <- ""
  # mismatch$temp[is.na(mismatch[,paste0(aaa[1])])] <- paste0("missing from hist")
  # mismatch$temp[is.na(mismatch[,paste0(aaa[2])])] <- paste0("missing from hist")
  # 
  # 
  # mismatch$temp <- ifelse((abs(mismatch[,paste0(aa[i], "_diff")])>0.05 | 
  #                            is.na(mismatch[,paste0(aa[i], "_diff")])), 
  #                         TRUE, FALSE)
  names(mismatch)[names(mismatch) == "temp"] <- paste0(aa[i], "_case")
}

mismatch00 <- mismatch
str00 <- paste0("There were ", nrow(mismatch), 
                " ", str0," estimates that changed from the ", 
                past_yr," and ", this_yr," datasets. ")

mismatch_missing <- mismatch[grepl(pattern = "missing from ", x = mismatch[, paste0(aa[1], "_case")]), ]
# table(mismatch_missing[, c("SPECIES_CODE", "YEAR", paste0(aa[i],"_case"))])
if (nrow(mismatch_missing) > 0) {
str00 <- paste0(str00, 
                "However, ", 
                ifelse(nrow(mismatch) != nrow(mismatch_missing), nrow(mismatch_missing), "all"),
                " observations were missing between datasets. ")
}
# table(mismatch_missing[, c("SPECIES_CODE", "YEAR")])

mismatch_rounding <- mismatch[grepl(pattern = "rounding error ", x = mismatch[, paste0(aa[1], "_case")]), ]
if (nrow(mismatch_rounding) > 0) {
str00 <- paste0(str00, 
                "", 
                ifelse(nrow(mismatch) != nrow(mismatch_rounding), nrow(mismatch_rounding), "all"),
                " differences were due to rounding errors (+-0.1 difference). ")
}

mismatch_actual <- mismatch[grepl(pattern = "actual change ", x = mismatch[, paste0(aa[1], "_case")]), ]
if (nrow(mismatch_actual) > 0) {
str00 <- paste0(str00, 
                "In total, ", 
                ifelse(nrow(mismatch) != nrow(mismatch_actual), nrow(mismatch_actual), "all"),
                " differences were actual changes. ")
}
mismatch <- mismatch_actual
# mismatch$case <- rowSums(is.na(mismatch[,which(grepl(pattern = "_case", x = names(mismatch)))]))
# mismatch$case <- rowSums(mismatch[,which(grepl(pattern = "_case", x = names(mismatch)))])
# mismatch$case <- ifelse(is.na(mismatch$case), TRUE, mismatch$case)


# mismatch <- mismatch[,which(grepl(pattern = "_case", x = names(mismatch)))]




# str00 <- paste0(a0$str00, 
#                 "With the new dataset, ", length(is.na(a[,paste0(aa[1], "_hist")])), 
#                 " enteries were added and ", length(is.na(a[,paste0(aa[1], "_prod")])), 
#                 " enteries were removed from the newest dataset. ")


mismatch <- mismatch %>% 
  # dplyr::filter(case>0) %>% 
  dplyr::select(dplyr::any_of(names(mismatch0)))

if (nrow(mismatch) == 0) {
  mismatch <- ""
} else {
  mismatch <- flextable::flextable(mismatch) %>% flextable::theme_vanilla()
}

return(list("a" = mismatch, 
            "a0" = mismatch00, 
            "str00" = str00))
}

```

```{r print_differences, echo=FALSE}
# str00 <- ""
# for (i in 1:length(all_mismatches)) {
#       str00 <- paste0(
#       str00,
#       "### ", names(all_mismatches)[i], "\n\n",
#       "\n\n",
#       kableExtra::kable(all_mismatches[i], row.names = FALSE, format = "html") %>%
#         kableExtra::kable_styling(bootstrap_options = "striped"),
#       "\n\n\n"
#     )
# }
# `r str00`
```

# Species

```{r}
a <- mismatch_spp %>% 
  dplyr::filter(SPECIES_NAME_diff) %>%
  dplyr::select(!dplyr::ends_with("_diff"))  %>%
  dplyr::select(SPECIES_CODE, dplyr::starts_with("SPECIES_NAME_")) 

str00 <- paste0("There were ", nrow(a), 
                " species scientific names that changed from the ", 
                past_yr," and ", this_yr," datasets. ")

a <- mismatch_spp_change
str00 <- paste0(str00, 
                "There were ", nrow(a), " species scientific taxonomics that changed across all of the surveys. ")

a0 <- table(mismatch_spp_change[, c("ACTION")]) %>% data.frame()
str00 <- paste0(str00, 
                "There were ", 
                paste0(a0$Freq, " instances of ", a0$Var1, collapse = " and "), 
                " species info tables that changed. ")
a0 <- table(mismatch_spp_change[, c("ACTION", "REASON")]) %>% data.frame()
str00 <- paste0(str00, 
                "These changes were changed because ", paste0(a0$REASON, collapse = ", "), " were identified. ")
```

```{r, eval = FALSE}
a <- mismatch_spp %>% 
  dplyr::filter(SPECIES_NAME_diff) %>%
  dplyr::select(!dplyr::ends_with("_diff"))  %>%
  dplyr::select(SPECIES_CODE, dplyr::starts_with("SPECIES_NAME_")) 

str00 <- paste0("These changes are now represented in the data. There were ", nrow(a), 
                " species scientific names that changed from the ", 
                past_yr," and ", this_yr," datasets. ")
```

```{r, eval = FALSE}
a <- mismatch_spp %>% 
  dplyr::filter(COMMON_NAME_diff) %>%
  dplyr::select(!dplyr::ends_with("_diff"))  %>%
  dplyr::select(SPECIES_CODE, dplyr::starts_with("common_")) 

str00 <- paste0("There were ", nrow(a), 
                " species common names that changed from the ", 
                past_yr," and ", this_yr," datasets. ")
```

`r (str00)` 

`r flextable::flextable(a) %>% flextable::theme_vanilla() `

# Catch-per-unit-effort

```{r changes-cpue}
a <- find_difference(mismatch = mismatch_cpue, 
                     str0 = "catch-per-unit-effort", 
                past_yr = past_yr, 
                this_yr = this_yr, 
                col = c("HAULJOIN", "SPECIES_CODE"))
```

`r (a$str00)` 

`r a$a`

# Biomass and Abundance

known cut offs for species because of confidence (ATF, Kam, skates; if n_haul of prod is NA and n_haul of hist is not 0 or NA)
unnecessary zero-filled for unobserved species in a survey (darkfin sculpin 21340; if n_haul of prod is NA and n_haul of hist is not 0)


```{r changes-bio}
a <- find_difference(mismatch = mismatch_biomass, 
                     str0 = "biomass and abundance", 
                past_yr = past_yr, 
                this_yr = this_yr, 
                col = c("SURVEY_DEFINITION_ID", "YEAR", "AREA_ID", "SPECIES_CODE"))
```

`r (a$str00)` 

<!-- `r a$a` -->

# Size Compositions

simply a data change: -9's were reconsidered as 'unsexed.' Tried to sex the fish but couldn't or were too small or data mixup. 

```{r changes-size}
a <- find_difference(mismatch = mismatch_sizecomp, 
                     str0 = "size compositions", 
                past_yr = past_yr, 
                this_yr = this_yr, 
                col = c("YEAR", "AREA_ID", "SPECIES_CODE", "SEX", "LENGTH_MM"))
```

`r (a$str00)` 

`r a$a`

# Age Compositions

```{r changes-age}
a <- find_difference(mismatch = mismatch_age, 
                     str0 = "age compositions", 
                past_yr = past_yr, 
                this_yr = this_yr, 
                col = c("YEAR", "AREA_ID", "SPECIES_CODE", "SEX", "LENGTH_MM", "AGE"))
```

`r (a$str00)` 

`r a$a`




