---
title: 'AFSC RACE Groundfish and Shellfish Survey Public Data'
description: |
 AFSC RACE Groundfish and Shellfish Survey Public Data
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```

This data contains all of the catch, environmental, and haul data from the fisheries-independent Groundfish and Shellfish Assessment Program surveys in the Bering Sea, Aleutian Islands, and Gulf of Alaska. This data is sought after by the general public, private entities, and NOAA partners alike, including tribal organizations, K-12 classrooms, academic institutions, for-profit groups, and non-profit groups. This data is compiled and approved once a year after each summer survey season and is approved for open access. 

# Collabrators and data users

Below are a few packages and products currently using this data. If you have developed a product, performed an analysis, or exhibited this data in any way, reach out so we can showcase your hard work. 

```{r citing-us, echo = FALSE}

str0 <- ""
 # df <- ""

if (tocTF == TRUE)  { # github markdown
  
str0 <- paste0(" - **[NOAA Fisheries Distribution Mapping and Analysis Portal](https://apps-st.fisheries.noaa.gov/dismap)**; *[NOAA Fisheries Office of Science and Technology](https://www.fisheries.noaa.gov/contact/office-science-and-technology)*
               
               
 - **[Pull data with python](https://pyafscgap.org/) and explore the [in-browser visualization tool](https://app.pyafscgap.org/'). Reference their [example Python notebook](https://mybinder.org/v2/gh/SchmidtDSE/afscgap/main?urlpath=/tree/index.ipynb)**; *[The Eric and Wendy Schmidt Center for Data Science and the Environment at UC Berkeley](https://dse.berkeley.edu/), including sam.pottinger@berkeley.edu, ccmartinez@berkeley.edu, gzarpellon@berkeley.edu, and kkoy@berkeley.edu.* ")
  
} else if (tocTF == FALSE)  { # html

 # DisMAP
df <- data.frame(    
    url = c("https://www.noaa.gov/sites/default/files/2022-03/noaa_emblem_logo-2022.png"), 
    Description = c("<a href = 'https://apps-st.fisheries.noaa.gov/dismap'>NOAA Fisheries Distribution Mapping and Analysis Portal</a>"), 
    Contacts = c("<a href = 'https://www.fisheries.noaa.gov/contact/office-science-and-technology'>NOAA Fisheries Office of Science and Technology</a>"))

 # DSE
df <- df %>% 
  dplyr::add_row(
    url = c("https://dse.berkeley.edu/sites/default/files/styles/openberkeley_image_full/public/general/dse_logostack.png"), 
                 Description = c("<a href = 'https://pyafscgap.org/'>Pull data with python</a> and explore the <a href = 'https://app.pyafscgap.org/'>in-browser visualization tool</a>. Reference their <a href = 'https://mybinder.org/v2/gh/SchmidtDSE/afscgap/main?urlpath=/tree/index.ipynb'>example Python notebook</a>"), 
                 Contacts = c("<a href = 'https://dse.berkeley.edu/'>The Eric and Wendy Schmidt Center for Data Science and the Environment at UC Berkeley</a>; <a href = 'mailto: sam.pottinger@berkeley.edu'>Sam Pottinger</a>, <a href = 'mailto: ccmartinez@berkeley.edu'>Ciera Martinez</a>, <a href = 'mailto: gzarpellon@berkeley.edu'>Giulia Zarpellon</a>, and <a href = 'mailto: kkoy@berkeley.edu'>Kevin Koy</a>"))

library(gt)
library(gtExtras)
library(purrr)
  
  df <- df %>%
  dplyr::select(url, 
                Description, Contacts) %>%
  dplyr::mutate(Contacts = map(Contacts, gt::html)) %>%
  dplyr::mutate(Description = map(Description, gt::html)) %>%
  gt::gt() %>%
  gtExtras::gt_img_rows(columns = url, img_source = "web") %>%
  gt::tab_options(data_row.padding = px(1)) %>%
  gt::cols_label(url = "     ") %>%
  cols_align(align = c("left"), columns = everything())
df
}
```

`r str0`

