---
title: 'Access public data using the API and R'
description: |
 Access public data using the API and R
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
library(dplyr)
library(magrittr)
```

# Connect to the API with R

An application programming interface (API) is a way for two or more computer programs to communicate with each other.

More information about how to amend API links can be found [here](https://docs.oracle.com/en/database/oracle/oracle-rest-data-services/22.3/books.html#AELIG90103/). Useful introductions to using APIs in `R` can be found [here](https://www.dataquest.io/blog/r-api-tutorial/). 

# API in action

## Ex. 1: Load the first 25 rows (default) of data

```{r test-0, echo = TRUE}
 # install.packages(c("httr", "jsonlite"))
library("httr")
library("jsonlite")
 # link to the API
api_link <- "https://apps-st.fisheries.noaa.gov/ods/foss/afsc_groundfish_survey/"

res <- httr::GET(url = api_link)
 # res # Test connection
data <- jsonlite::fromJSON(base::rawToChar(res$content))
 # names(data)
flextable::flextable(head(data$items, 3)) 
```

## Ex. 2: Load the first 10000 rows of data

```{r test-1}
res <- httr::GET(url = paste0(api_link, "?offset=0&limit=10000"))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
dim(data$items)
flextable::flextable(data$items)
```

## Ex. 3: Filter by Year

Show all the data greater than the year 2020. 

```{r test-2, eval = FALSE, echo = FALSE}
res <- httr::GET(url = paste0(api_link, '?q={"year":{"$gt":2020}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
flextable::flextable(data$items[1:3, c("year", "srvy", "stratum", "species_code", "cpue_kgkm2")])
```

## Ex. 4: Filter by Species name

Show all the data where the product name contains pollock Please note that here the word pollock is case sensitive.

The notation for finding a string is to use % around it. Since % is a reserved character in a URL, you have to replace `%` with `%25`. 

```{r test-3}
res <- httr::GET(url = paste0(api_link, '?q={"common_name":{"$like":"%25pollock%25"}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
flextable::flextable(data$items[1:3, c("year", "srvy", "stratum", "species_code", "cpue_kgkm2")])
```

## Ex. 4: Combination of year and name filters

Show all the data where years > 2020 and the product name contains pollock

```{r test-4, eval = FALSE, echo = FALSE}
res <- httr::GET(url = paste0(api_link, '?q={"year":{"$gt":2020},"name":{"$like":"%25pollock%25"}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
flextable::flextable(data$items[1:3, c("year", "srvy", "stratum", "species_code", "cpue_kgkm2")])
```

## Ex. 5: Combination of year, srvy, stratum

Show all the data where year = 1989, srvy = "EBS", and stratum is not equal to 81

```{r test-5, eval = FALSE, echo = TRUE}
res <- httr::GET(url = paste0(api_link, '?q={"year":1989,"srvy":"EBS","stratum":{"$ne":"81"}}'))
data <- jsonlite::fromJSON(base::rawToChar(res$content))
flextable::flextable(data$items[1:3, c("year", "srvy", "stratum", "species_code", "cpue_kgkm2")])
```

## Ex. 6: Visualize CPUE data in distribution map

Pacific cod catch-per-unit-effort estimates for NBS in 2021 and map constructed using [`akgfmaps`](https://github.com/afsc-gap-products/akgfmaps). 

```{r test-6, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE}
# res <- httr::GET(
#   url = paste0(api_link, "?offset=0&limit=10000"), 
#   query = list(year = 2021, srvy = "EBS", species_code = 30060))
res <- httr::GET(url = paste0(api_link, '?q={"year":2021,"srvy":"NBS","species_code":21720}'))
data_catch <- jsonlite::fromJSON(base::rawToChar(res$content))$items %>% 
  dplyr::select(stratum, station, cpue_kgkm2) 

# zero-fill data (imperfectly, but effective for this example)
res <- httr::GET(url = paste0(api_link, '?q={"year":2021,"srvy":"NBS"}offset=0&limit=10000'))
data_haul <- jsonlite::fromJSON(base::rawToChar(res$content))$items %>% 
  dplyr::select(stratum, station, latitude_dd, longitude_dd) %>%
  dplyr::distinct()

data <- dplyr::left_join(data_haul, data_catch) %>% 
  dplyr::mutate(cpue_kgkm2 = ifelse(is.na(cpue_kgkm2), 0, cpue_kgkm2), 
                dplyr::across(dplyr::everything(), as.numeric)) 

# devtools::install_github("afsc-gap-products/akgfmaps", build_vignettes = TRUE)
library(akgfmaps)

figure <- akgfmaps::make_idw_map(
  CPUE_KGHA = data$cpue_kgkm2, # calculates the same, regardless of units.  
  LATITUDE = data$latitude_dd, 
  LONGITUDE = data$longitude_dd, 
  region = "bs.north", # Predefined EBS area
  set.breaks = "jenks", # Gets Jenks breaks from classint::classIntervals()
  in.crs = "+proj=longlat", # Set input coordinate reference system
  out.crs = "EPSG:3338", # Set output coordinate reference system
  grid.cell = c(20000, 20000), # 20x20km grid
  key.title = "Pacific Ocean perch") # Include in the legend title

flextable::flextable(data[1:3,])
figure$plot + 
  ggplot2::guides(fill=guide_legend(title = "Pacific cod\nCPUE (kg/km2)"))
```


```{r child-footer, child=here::here("docs","footer.Rmd")}
```
