---
title: 'Access public data using R in Oracle'
description: |
 Access public data using R in Oracle (AFSC only)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, error = FALSE, message = FALSE)
```

## Access data via Oracle

If the user has access to the AFSC `Oracle` database, the user can use `SQL developer` to view and pull the FOSS public data directly from the `RACEBASE_FOSS` `Oracle` schema. 

### Connect to Oracle from R

Many users will want to access the data from `Oracle` using `R`. The user will need to install the `RODBC` `R` package and ask OFIS (IT) connect `R` to `Oracle`. Then, use the following code in `R` to establish a connection from `R` to `Oracle`: 

Here, the user can write in their username and password directly into the `RODBC` connect function. Never save usernames or passwords in scripts that may be intentionally or unintentionally shared with others. If no username and password is entered in the function, pop-ups will appear on the screen asking for the username and password. 

```{r oracle-connect-2, echo = TRUE, eval = FALSE}
 #' Define RODBC connection to ORACLE
 #'
 #' @param schema default = 'AFSC'. 
 #'
 #' @return oracle channel connection
 #' @export
 #'
 #' @examples
 #' # Not run
 #' # channel <- oracle_connect()
oracle_connect <- function(
    schema='AFSC', 
    username = NULL, 
    passowrd = NULL){(echo=FALSE)
  
  library("RODBC")
  library("getPass")
  if (is.null(username)) {
    username <- getPass(msg = "Enter your ORACLE Username: ")
  }
  if (is.null(password)) {
    password <- getPass(msg = "Enter your ORACLE Password: ")
  }
  channel  <- RODBC::odbcConnect(
    paste(schema),
    paste(username),
    paste(password), 
    believeNRows=FALSE)
  return(channel)
}

channel <- oracle_connect()
```

### Ex. 1: Select all data

Once connected, pull and save (if needed) the table into the `R` environment. 

```{r oracle-data-pull, echo = TRUE, eval = FALSE}
# Pull table from oracle into R environment
a <- RODBC::sqlQuery(channel, "SELECT * FROM RACEBASE_FOSS.FOSS_CPUE_ZEROFILLED")
# Save table to local directory
write.csv(x = a, file = "RACEBASE_FOSS-FOSS_CPUE_ZEROFILLED.csv")
```

### Ex. 2: Subset data

To pull a small subset of the data (especially since files like `RACEBASE_FOSS.FOSS_CPUE_ZEROFILLED` are so big), use a variation of the following code. Here, we are pulling EBS Pacific cod from 2010 - 2021: 

```{r oracle-data-pull-subset, echo = TRUE, eval = FALSE}
 # Pull data
a <- RODBC::sqlQuery(channel, "SELECT * FROM RACEBASE_FOSS.FOSS_CPUE_ZEROFILLED 
WHERE SRVY = 'EBS' 
AND COMMON_NAME = 'Pacific cod' 
AND YEAR >= 2010 
AND YEAR < 2021")
 # Save table to local directory
write.csv(x = a, file = "RACEBASE_FOSS-FOSS_CPUE_ZEROFILLED-ebs_pcod_2010-2020.csv")
```


### Join catch and haul data

If this file is too large, you can join catch and haul data. Pull the `RACEBASE_FOSS.JOIN_FOSS_CPUE_CATCH` and  `RACEBASE_FOSS.JOIN_FOSS_CPUE_HAUL` tables, which are much smaller, and combine the table locally yourself. 


```{r oracle-data-size-join-catch, echo = FALSE, eval = FALSE}

# <!-- `r paste(readLines(con = paste0(dir_out, "JOIN_FOSS_CPUE_metadata_table.txt")), collapse="\n") ` -->

str0 <- paste0(
  "RACEBASE_FOSS.JOIN_FOSS_CPUE_CATCH: ", "
  rows: ", nrow(JOIN_FOSS_CPUE_CATCH), "
  cols: ", ncol(JOIN_FOSS_CPUE_CATCH), "
  ", round(file.size(paste0(dir_out, "JOIN_FOSS_CPUE_CATCH.csv"))/1e+9, 3), " GB") 
cat(str0)
```

```{r oracle-data-size-join-haul, echo = FALSE, eval = FALSE}
str0 <- paste0(
  "RACEBASE_FOSS.JOIN_FOSS_CPUE_HAUL: ", "
  rows: ", nrow(JOIN_FOSS_CPUE_HAUL), "
  cols: ", ncol(JOIN_FOSS_CPUE_HAUL), "
  ", round(file.size(paste0(dir_out, "JOIN_FOSS_CPUE_HAUL.csv"))/1e+9, 3), " GB") 
cat(str0)
```

To join these tables in Oracle, you may use a variant of the following code: 

```{sql, eval = FALSE}
SELECT * FROM RACEBASE_FOSS.JOIN_FOSS_CPUE_HAUL
FULL JOIN RACEBASE_FOSS.JOIN_FOSS_CPUE_CATCH
ON RACEBASE_FOSS.JOIN_FOSS_CPUE_HAUL.HAULJOIN = RACEBASE_FOSS.JOIN_FOSS_CPUE_CATCH.HAULJOIN;
```


```{=html}
<script>
function myFunction() {
var input, filter, table, tr, td, i, txtValue;
input = document.getElementById("myInput");
filter = input.value.toUpperCase();
table = document.getElementById("myTable");
tr = table.getElementsByTagName("tr");
for (i = 0; i < tr.length; i++) {
td = tr[i].getElementsByTagName("td")[1];
if (td) {
txtValue = td.textContent || td.innerText;
if (txtValue.toUpperCase().indexOf(filter) > -1) {
tr[i].style.display = "";
} else {
tr[i].style.display = "none";
}
}       
}
}
</script>
```

